Making Sequential Programs Deterministic

### 2019-03-28 design.tex
Edits in section 2.2

    $\infer{\tau_1 \sqsubseteq \tau_3}{\tau_1 \sqsubseteq \tau_2 \quad \tau_2 \sqsubseteq \tau_3}
    \quad\quad
    \infer[\textsc{subtyping}]{\Gamma \vdash x : \tau_2}{\Gamma \vdash x : \tau_1 & \tau_1 \sqsubseteq \tau_2}$

@@

    $\infer[\textsc{method call}]{\|method|(\overline{a_i}) : \tau_r}{\tau_r \  \|method|(\overline{\tau_i \  p_i}) & \overline{a_i : \tau_i}}$

@@

    $\infer[\textsc{invariant type arg}]{C_1\angles{\overline{\tau_i}}
      \sqsubseteq C_2\angles{\overline{\tau_i}}}{C_1 \sqsubseteq C_2}$

@@

\begin{wrapfigure}{R}{0.3\textwidth}
    \begin{center}
        \includegraphics[scale=0.5]{determinism}
    \end{center}
    \caption{Determinism type qualifier hierarchy}
    \label{fig:determinism-hierarchy}
\todo{Drop the at-signs (@).  The text is weirdly small for the box.
  Probably just redraw this.}
\end{wrapfigure}

@@

A map is a dictionary or associative array, such as a hash table.
Our type system largely treats a map as a collection of key--value pairs.

@@

@@

\todo{Note that the \textsc{ordernondet} versions also handle \<Det>,
  because $\<Det> \sqsubseteq \<OrderNonDet>$.}

\begin{figure}
    $\infer[\textsc{any type}]{\wellformed{\kappa \  \beta}}{\kappa : \|Det / NonDet|}$

@@

    $\infer[\textsc{ordernondet collection}]{\wellformed{\kappa_c \  \beta_c \angles{\kappa_e \ \beta_e}}}{\kappa_c : \|OrderNonDet/Det| & \kappa_e \sqsubseteq \kappa_c & \beta_c \sqsubseteq \<Collection>}$

@@

    $\infer[\textsc{nondet collection}]{\wellformed{\|NonDet| \ \beta_c\angles{\|NonDet| \ \beta_e}}}{\beta_c \sqsubseteq \<Collection>}$

@@

%      $\infer[\textsc{ordernondet array}]{\wellformed{\kappa_e \ \beta_e \ \  \kappa_a []}}{\kappa_a : \|OrderNonDet/Det| & \kappa_e \sqsubseteq \kappa_a }$
%     
%     \bigskip
%     
%     $\infer[\textsc{nondet array}]{\wellformed{\|NonDet| \ \beta_e \ \  \|NonDet| []}}{}$
%     
%     \bigskip

    $\infer[\textsc{ordernondet map}]{\wellformed{\kappa_m \  \beta_m \angles{\kappa_k \ \beta_k, \kappa_v \ \beta_v}}}{\kappa_m: \|OrderNonDet/Det| & \kappa_k \sqsubseteq \kappa_m & \kappa_v \sqsubseteq \kappa_m & \beta_m \sqsubseteq \<Map>}$

@@

    $\infer[\textsc{nondet map}]{\wellformed{\|NonDet| \ \beta_m\angles{\|NonDet| \ \beta_k, \|NonDet| \ \beta_v}}}{\beta_m \sqsubseteq \<Map>}$

    \caption{Type well formedness}
\todo{The \textsc{any type} rule is not correct, because not every type is
  well-formed.  It needs an antecedent stating that $\beta$ is not a
  collection or map.}

@@

\Cref{type-validity} shows the type validity\todo{The caption says
  ``type well formedness''.  Choose one term and use it consistently
  throughout.}
rules of our type system. A type $\tau$ is represented as $\kappa \ \beta$
where $\kappa$ is a type qualifier (that is, \<NonDet>, \<OrderNonDet>, or
\<Det>) and $\beta$ is a Java basetype.
As stated in the rule $\textsc{any type}$, the types \<Det> and \<NonDet>
can be written on any Java basetype.\todo{As noted in figure caption, that's
  not true, in some contexts such as a type argument.}

The determinism type qualifier on a collection must be a supertype or equal to
its element type qualifier (Rule $\textsc{ordernondet collection}$ in \cref{fig-determinism-collections}).
If the Collection is \<NonDet>, then the type argument may not be
\<Det> or \<OrderNonDet> (Rule $\textsc{nondet collection}$ in \cref{type-validity}). Although such types could exist, they are
disallowed to prevent the following bug:

@@

\smallskip
\noindent
\begin{minipage}{.48\textwidth}
Some examples of valid types are:
% Using "~" instead of "\ " does not work within \codeid :-(
\begin{smaller}
\begin{Verbatim}
@NonDet      List<@NonDet      Integer>
@OrderNonDet List<@OrderNonDet Set<...>>
@OrderNonDet List<@Det         Integer>
@Det         List<@Det         Integer>

\end{Verbatim}
\end{smaller}
\end{minipage}
\hfill
\begin{minipage}{.46\textwidth}
These types are invalid:
\begin{smaller}
\begin{Verbatim}
@NonDet      List<@OrderNonDet Set<...>>
@NonDet      List<@Det         Integer>
@OrderNonDet List<@NonDet      Integer>
@Det         List<@NonDet      Integer>
@Det         List<@OrderNonDet Set<...>>
\end{Verbatim}
\end{smaller}
\end{minipage}
\smallskip

\begin{figure}
    \centering
    \begin{tabular}{|l|l|l|l|l|}
        \cline{3-5}
        \multicolumn{2}{c|}{~}  &  \multicolumn{3}{c|}{element type} \\ \cline{3-5}
        \multicolumn{2}{c|}{~}  & NonDet     & OrderNonDet & Det \\ \hline
        & NonDet      &   valid    &  invalid    & invalid  \\ \cline{2-5}
        Collection    & OrderNonDet &   invalid  &  valid      & valid  \\ \cline{2-5}
        & Det         &   invalid  &  invalid    & valid      \\ \hline
    \end{tabular}
    \caption{Valid Collection declarations.  The Collection's type qualifier
        must be a supertype or equal to the element type.}
    \label{fig-determinism-collections}
\end{figure}

@@

Our type system supports three types of polymorphism:  type
polymorphism, basetype polymorphism, and qualifier polymorphism.
\todo{Give an example for each.}
\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forall \kappa . \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}

@@

\begin{figure}
    $\infer[\textsc{poly resolution}]{\pi_r \ \beta_r\ \<m>(\overline{\pi_i \ \beta_i})\{\ \}}{\forall \kappa \in \{\|NonDet, OrderNonDet, Det|\} : \kappa \ \beta_r\ \<m>(\overline{K \ \beta_i})\{\ \} & \pi_r, \pi_i \in \{\|PolyDet|, \|PolyDet|\uparrow, \|PolyDet|\downarrow    \}}$
    \caption{Polymorphic resolution rule}
    \todo{I'm not sure what this is doing.  What code does it reject?}
    \todo{This suggests that \PolyDetUp is a qualifier, rather than
      $\uparrow$ being an operator on qualifiers.  We should clarify which
      it is.}
    \label{fig-poly-resolution}
\end{figure}

@@

In \theDeterminismChecker, we chose \<PolyDet> as the default for all 
method parameters and return types including constructors.
The rationale for this design choice was to not make our type checker too restrictive by making everything \<Det>.  This implies that
\theDeterminismChecker still allows some non-determinism as long as it doesn't break any determinism specifications explicitly written by the programmer.

@@

\TheDeterminismChecker has additional type refinement rules for the sorting
methods such as \<List.sort>, \<Collections.sort>, and \<Arrays.sort>:
 as shown in \cref{fig-sorting}.
\begin{figure}
%     $\infer[\textsc{list sort}]{\Gamma \vdash x : \|Det \ List|\angles{\|Det| \ \beta}}{\Gamma \vdash x : \|OrderNonDet \ List|\angles{\|Det| \ \beta} & \|x.sort()|}$
%     
%     \bigskip

    $\infer[\textsc{sort}]{\Gamma \vdash x : \|Det \ List|\angles{\|Det| \ \beta}}{\Gamma \vdash x : \|OrderNonDet \ List|\angles{\|Det| \ \beta} & \|Collections.sort(x)|}$

%     \bigskip
%     
%     $\infer[\textsc{arrays sort}]{\Gamma \vdash x : \|Det \ \beta \ Det[]| }{\Gamma \vdash x : \|Det \ \beta \ OrderNonDet[]| & \|Arrays.sort( \ x)|}$

@@

\begin{figure}
    $\infer[\textsc{hashmap creation}]{\|new \ \kappa\ HashMap()|}{\kappa = \|OrderDet/NonDet|}$

@@

    $\infer[\textsc{treemap creation}]{\|new \ \kappa\ TreeMap()|}{\kappa = \|Det/NonDet|}$

@@

    $\infer[\textsc{hashset creation}]{\|new \ \kappa\ HashSet()|}{\kappa = \|OrderDet/NonDet|}$

@@

    $\infer[\textsc{treeset creation}]{\|new \ \kappa\ TreeSet()|}{\kappa = \|Det/NonDet|}$
    \caption{\<Map> and \<Set> creation rules}
    \label{fig-creation-rules}
\end{figure}

@@

\subsection{Determinism checker specific rules for precision}\label{precision}
We annotated the \codeid{equals()} method of the \<Set> class in the JDK as
\begin{verbatim}
PolyDet$\uparrow$ boolean equals(PolyDet Set<E> this, PolyDet Object o)
\end{verbatim}
While this is sound, it is imprecise in the case of \codeid{OrderNonDet Sets}. We added the following rule shown in \cref{fig-precision-rules} to our checker to improve precision.
\begin{figure}
    \todo{add line break}
    $\infer[\textsc{set precision}]{\Gamma \vdash \|z : Det|}{\Gamma \vdash \|x| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{x} \ \beta_{x}}, \|y| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{y} \ \beta_{y}} & \|z = x.equals(y)| & \angles{ \kappa_{x} \ \beta_{x}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } & \angles{ \kappa_{y} \ \beta_{y}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } }$

@@

    $\infer[\textsc{map precision}]{\Gamma \vdash \|z : Det|}{\Gamma \vdash \|x : OrderNonDet \ Map\angles{ Det \ \beta_k, \kappa_v\ \beta_v }, y : Det \ \beta| & \|z = x.get(y)|}$

>>>>

    $\infer{\tau_1 \sqsubseteq \tau_3}{\tau_1 \sqsubseteq \tau_2 \quad \tau_2 \sqsubseteq \tau_3}
    \quad\quad
    \infer[\rulename{subtyping}]{\Gamma \vdash x : \tau_2}{\Gamma \vdash x : \tau_1 & \tau_1 \sqsubseteq \tau_2}$

@@

    $\infer[\rulename{method call}]{\|method|(\overline{a_i}) : \tau_r}{\tau_r \  \|method|(\overline{\tau_i \  p_i}) & \overline{a_i : \tau_i}}$

@@

    $\infer[\rulename{invariant type arg}]{C_1\angles{\overline{\tau_i}}
      \sqsubseteq C_2\angles{\overline{\tau_i}}}{C_1 \sqsubseteq C_2}$

@@

\begin{wrapfigure}{R}{0.3\textwidth}
    \begin{center}
        \includegraphics[scale=0.5]{determinism}
    \end{center}
    \caption{Determinism type qualifier hierarchy}
    \label{fig:determinism-hierarchy}
\todo{Drop the at-signs (@).}
\todo{The text is weirdly small for the box.
  Probably redraw this.}
\end{wrapfigure}

@@

A map is a dictionary or associative array, such as a hash table.
\OurTypeSystem largely treats a map as a collection of key--value pairs.

@@

\todo{A type $\tau$ is represented as $\kappa \ \beta$
where $\kappa$ is a type qualifier (that is, \<NonDet>, \<OrderNonDet>, or
\<Det>) and $\beta$ is a Java basetype.}

@@

\todo{Note that the \rulename{ordernondet} versions also handle \<Det>,
  because $\<Det> \sqsubseteq \<OrderNonDet>$.}

\begin{figure}
    $\infer[\rulename{any type}]{\wellformed{\kappa \  \beta}}{\kappa : \|Det / NonDet|}$

@@

    $\infer[\rulename{ordernondet collection}]{\wellformed{\kappa_c \  \beta_c \angles{\kappa_e \ \beta_e}}}{\kappa_c : \|OrderNonDet/Det| & \kappa_e \sqsubseteq \kappa_c & \beta_c \sqsubseteq \<Collection>}$

@@

    $\infer[\rulename{nondet collection}]{\wellformed{\|NonDet| \ \beta_c\angles{\|NonDet| \ \beta_e}}}{\beta_c \sqsubseteq \<Collection>}$

@@

%      $\infer[\rulename{ordernondet array}]{\wellformed{\kappa_e \ \beta_e \ \  \kappa_a []}}{\kappa_a : \|OrderNonDet/Det| & \kappa_e \sqsubseteq \kappa_a }$
%     
%     \bigskip
%     
%     $\infer[\rulename{nondet array}]{\wellformed{\|NonDet| \ \beta_e \ \  \|NonDet| []}}{}$
%     
%     \bigskip

    $\infer[\rulename{ordernondet map}]{\wellformed{\kappa_m \  \beta_m \angles{\kappa_k \ \beta_k, \kappa_v \ \beta_v}}}{\kappa_m: \|OrderNonDet/Det| & \kappa_k \sqsubseteq \kappa_m & \kappa_v \sqsubseteq \kappa_m & \beta_m \sqsubseteq \<Map>}$

@@

    $\infer[\rulename{nondet map}]{\wellformed{\|NonDet| \ \beta_m\angles{\|NonDet| \ \beta_k, \|NonDet| \ \beta_v}}}{\beta_m \sqsubseteq \<Map>}$

    \caption{Type well formedness}
\todo{The \rulename{any type} rule is not correct, because not every type is
  well-formed.  It needs an antecedent stating that $\beta$ is not a
  collection or map.}

@@

\Cref{type-validity} shows the type validity\todo{The caption says
  ``type well formedness''.  Choose one term and use it consistently
  throughout.}
rules of \ourTypeSystem. 
As stated in the rule \rulename{any type}, the types \<Det> and \<NonDet>
can be written on any Java basetype.\todo{As noted in figure caption, that's
  not true, in some contexts such as a type argument.}

\begin{figure}
    \centering
    \begin{tabular}{|l|l|l|l|l|}
        \cline{3-5}
        \multicolumn{2}{c|}{~}  &  \multicolumn{3}{c|}{element type} \\ \cline{3-5}
        \multicolumn{2}{c|}{~}  & NonDet     & OrderNonDet & Det \\ \hline
        & NonDet      &   valid    &  invalid    & invalid  \\ \cline{2-5}
        Collection    & OrderNonDet &   invalid  &  valid      & valid  \\ \cline{2-5}
        & Det         &   invalid  &  invalid    & valid      \\ \hline
    \end{tabular}
    \caption{Valid Collection declarations.  The Collection's type qualifier
        must be a supertype or equal to the element type.}
    \label{fig:determinism-collections}
\end{figure}

The element type of a \<NonDet> collection must be \<NonDet>.
Otherwise, the type qualifier on a collection's element must be a subtype or equal to
the collection's type qualifier (\rulename{collection}\ rules of
\cref{type-validity}).
\Cref{fig:determinism-collections} restates its restrictions, and here are examples.

\smallskip
\noindent
\begin{minipage}{.48\textwidth}
Some examples of valid types are:
% Using "~" instead of "\ " does not work within \codeid :-(
\begin{smaller}
\begin{Verbatim}[commandchars=\\\{\}]
NonDet      \lptype{List}{NonDet      Integer}
OrderNonDet \lptype{List}{OrderNonDet \lptype{Set}{...}}
OrderNonDet \lptype{List}{Det         Integer}
Det         \lptype{List}{Det         Integer}

\end{Verbatim}
\end{smaller}
\end{minipage}
\hfill
\begin{minipage}{.46\textwidth}
These types are invalid:
\begin{smaller}
\begin{Verbatim}[commandchars=\\\{\}]
NonDet      \lptype{List}{OrderNonDet \lptype{Set}{...}}
NonDet      \lptype{List}{Det         Integer}
OrderNonDet \lptype{List}{NonDet      Integer}
Det         \lptype{List}{NonDet      Integer}
Det         \lptype{List}{OrderNonDet \lptype{Set}{...}}
\end{Verbatim}
\end{smaller}
\end{minipage}
\smallskip

 Although such types could exist, they are
disallowed to prevent the following bug:

@@

@@

\OurTypeSystem supports three types of polymorphism:  type
polymorphism, basetype polymorphism, and qualifier polymorphism.
\todo{Give an example for each.}
\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forall \kappa . \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}

@@

\begin{figure}
    $\infer[\rulename{poly resolution}]{\pi_r \ \beta_r\ \<m>(\overline{\pi_i \ \beta_i})\{\ \}}{\forall \kappa \in \{\|NonDet, OrderNonDet, Det|\} : \kappa \ \beta_r\ \<m>(\overline{K \ \beta_i})\{\ \} & \pi_r, \pi_i \in \{\|PolyDet|, \|PolyDet|\uparrow, \|PolyDet|\downarrow    \}}$
    \caption{Polymorphic resolution rule}
    \todo{I'm not sure what this is doing.  What code does it reject?}
    \todo{This suggests that \PolyDetUp is a qualifier, rather than
      $\uparrow$ being an operator on qualifiers.  We should clarify which
      it is.}
    \label{fig-poly-resolution}
\end{figure}

@@

In \theDeterminismChecker, we chose \<PolyDet> as the default for all 
method parameters and return types including constructors.
The rationale for this design choice was to not make \theDeterminismChecker too restrictive by making everything \<Det>.  This implies that
\theDeterminismChecker still allows some non-determinism as long as it doesn't break any determinism specifications explicitly written by the programmer.

@@

\TheDeterminismChecker has additional type refinement rules for the sorting
methods such as \<List.sort>, \<Collections.sort>, and \<Arrays.sort>:
 as shown in \cref{fig-sorting}.
\begin{figure}
%     $\infer[\rulename{list sort}]{\Gamma \vdash x : \|Det \ List|\angles{\|Det| \ \beta}}{\Gamma \vdash x : \|OrderNonDet \ List|\angles{\|Det| \ \beta} & \|x.sort()|}$
%     
%     \bigskip

    $\infer[\rulename{sort}]{\Gamma \vdash x : \|Det \ List|\angles{\|Det| \ \beta}}{\Gamma \vdash x : \|OrderNonDet \ List|\angles{\|Det| \ \beta} & \|Collections.sort(x)|}$

%     \bigskip
%     
%     $\infer[\rulename{arrays sort}]{\Gamma \vdash x : \|Det \ \beta \ Det[]| }{\Gamma \vdash x : \|Det \ \beta \ OrderNonDet[]| & \|Arrays.sort( \ x)|}$

@@

\begin{figure}
    $\infer[\rulename{hashmap creation}]{\|new \ \kappa\ HashMap()|}{\kappa = \|OrderDet/NonDet|}$

@@

    $\infer[\rulename{treemap creation}]{\|new \ \kappa\ TreeMap()|}{\kappa = \|Det/NonDet|}$

@@

    $\infer[\rulename{hashset creation}]{\|new \ \kappa\ HashSet()|}{\kappa = \|OrderDet/NonDet|}$

@@

    $\infer[\rulename{treeset creation}]{\|new \ \kappa\ TreeSet()|}{\kappa = \|Det/NonDet|}$
    \caption{\<Map> and \<Set> creation rules}
    \label{fig-creation-rules}
\end{figure}

@@

\subsection{Determinism checker specific rules for precision}\label{precision}
We annotated the \codeid{equals()} method of the \<Set> class in the JDK as
\begin{verbatim}
PolyDet$\uparrow$ boolean equals(PolyDet Set<E> this, PolyDet Object o)
\end{verbatim}
While this is sound, it is imprecise in the case of \codeid{OrderNonDet Sets}. We added the following rule shown in \cref{fig-precision-rules} to our checker to improve precision.
\begin{figure}
    \todo{add line break}
    $\infer[\rulename{set precision}]{\Gamma \vdash \|z : Det|}{\Gamma \vdash \|x| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{x} \ \beta_{x}}, \|y| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{y} \ \beta_{y}} & \|z = x.equals(y)| & \angles{ \kappa_{x} \ \beta_{x}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } & \angles{ \kappa_{y} \ \beta_{y}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } }$

@@

    $\infer[\rulename{map precision}]{\Gamma \vdash \|z : Det|}{\Gamma \vdash \|x : OrderNonDet \ Map\angles{ Det \ \beta_k, \kappa_v\ \beta_v }, y : Det \ \beta| & \|z = x.get(y)|}$

### 2019-03-28 introduction.tex
Edits in section 2.0
Edits in section 2.2

Our analysis uses three main abstractions, or approximations to run-time values:
\begin{description}
\item[\<Det>] represents values that will be the same across executions.
\item[\<OrderNonDet>] represents collections that are guaranteed to contain the same elements albeit in possibly
different iteration order.
\item[\<NonDet>] represents values that
might differ from run to run.
\end{description}
\noindent
Programmers can write these to specify their program's behavior.
Our type system contains a variety of other features that increase
expressiveness and reduce false positive alarms.\todo{reference a section}

@@

  \item In a case study, we annotated several libraries, including the
    relevant parts of the JDK (Java's standard library).  This provides a
    formal, machine-readable specification for the libraries, and it
    demonstrates the richness of our type system.

@@

>>>>

Our analysis uses three main abstractions, or approximations to run-time values:
\begin{description}
\item[\<Det>] represents values that will be the same across executions.
\item[\<OrderNonDet>] represents collections that are guaranteed to contain the same elements albeit in possibly
different iteration order.
\item[\<NonDet>] represents values that
might differ from run to run.
\end{description}
\noindent
Programmers can write these to specify their program's behavior.
\OurTypeSystem contains a variety of other features that increase
expressiveness and reduce false positive alarms.\todo{reference a section}

@@

  \item In a case study, we annotated several libraries, including the
    relevant parts of the JDK (Java's standard library).  This provides a
    formal, machine-readable specification for the libraries, and it
    demonstrates the richness of \ourTypeSystem.

@@

\todo{It is essential that the introduction includes an example real-world
  defect that \theDeterminismChecker found.}
